---
title: "R preprocessing and plotting"
author: "Sophia Kleist Karlson"
date: "1 dec 2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load necesary packages through pacman
pacman::p_load(tidyverse, stringr, dplyr, tidyr, ggplot2)

# set working directory
setwd("~/Social Transmission Study/Analysis of drawings/")

# RStudio version
RStudio.Version()
```


NB:
BEFORE RUNNING THIS SCRIPT, RUN THE PYTHON IMAGE PROCESSING SCRIPT, WHICH CREATES THE COMPLEXITY_COMPARISON.CSV THAT IS USED BELOW



Comparing complexity of orignal and blurred images
```{r}
# read the cleaned_data.csv that was made in the R_preprocessing script
all_data <- read.csv("data/csv_files/cleaned_data_small.csv")
all_data$X <- NULL

# read complexity data and delete the first unnecessary column
complexity_data <- read.csv("data/csv_files/complexity_comparison.csv")
complexity_data$X <- NULL


# check class of complexity measures - they need to be numeric
class(complexity_data$Complexity_original)
class(complexity_data$Complexity_convolution)

# make numeric
complexity_data$Complexity_original <- as.numeric(complexity_data$Complexity_original)
complexity_data$Complexity_convolution <- as.numeric(complexity_data$Complexity_convolution)


# make a column with the rato of complexity between the blurred and original images
complexity_data <- complexity_data %>% mutate(ratio = Complexity_convolution/Complexity_original)
range(complexity_data$ratio)


# correlation between the ratio and the comlexity of orignals
cor(complexity_data$ratio, complexity_data$Complexity_original) #-0.1768 - pretty weak correlation, but still there
cor(complexity_data$ratio, complexity_data$Complexity_convolution) # -0.1259 - even weaker
```



Merge complexity dataframe with all_data
```{r}
all_data_2 <- merge(all_data, complexity_data)

all_data_2$Complexity <- all_data_2$Complexity_convolution

all_data_2$Complexity_convolution <- NULL


#write csv file with complexity scores
write.csv(all_data_2, "data/csv_files/data_w_complexity.csv")
```



Read conventionality scores from all_conventionality.csv
```{r}
all_conventionality <- read.csv("data/csv_files/all_conventionality.csv")
all_conventionality$X <- NULL
```



Make extra rows for source images for each chain, so they can function as generation 0 - we use all_data_2
```{r}
# read data (if you have cleaned the environment after the last chunk)
#all_data_2 <- read.csv("data/csv_files/data_w_complexity.csv")
#all_data_2$X <- NULL

# make new dataframe from all_data_2
all_data_w_source <- all_data_2

# add 240 empty rows to be filled with source images in the beginning of each chain
all_data_w_source[nrow(all_data_w_source)+240,] <- NA


# check classes and make them into numeric if necesarry
class(all_data_w_source$Generation)
class(all_data_w_source$Condition)
class(all_data_w_source$Chain)
class(all_data_w_source$Source_image)
class(all_data_w_source$Complexity)

all_data_w_source$Generation <- as.numeric(all_data_w_source$Generation)
all_data_w_source$Condition <- as.numeric(all_data_w_source$Condition)
all_data_w_source$Chain <- as.numeric(all_data_w_source$Chain)
all_data_w_source$Source_image <- as.numeric(all_data_w_source$Source_image)


# make every generation 1 higher
all_data_w_source <- all_data_w_source %>% mutate(Generation = Generation +1)



# read csv with complexity of source images
source_comp <- read.csv("data/source_images/complexity/complexity_comparison_source.csv")
source_comp$X <- NULL

# add generation column which is 0
source_comp$Generation <- rep(0,12)

# add source image column, which is just the drawing id without "stim_"
source_comp <- source_comp %>% mutate(Source_image = str_replace_all(Drawing_ID, 'stim_', ''))

# make the source_image column numeric
class(source_comp$Source_image)
source_comp$Source_image <- as.numeric(source_comp$Source_image)

# make drawing id into character for both datasets
class(source_comp$Drawing_ID)
source_comp$Drawing_ID <- as.character(source_comp$Drawing_ID)

class(all_data_w_source$Drawing_ID)
all_data_w_source$Drawing_ID <- as.character(all_data_w_source$Drawing_ID)


# add columns from source_comp to all_data_w_source
all_data_w_source[1681:1692,1] <- source_comp$Drawing_ID
all_data_w_source[1681:1692,4] <- source_comp$Generation
all_data_w_source[1681:1692,6] <- source_comp$Source_image
all_data_w_source[1681:1692,8] <- source_comp$Complexity_original
all_data_w_source[1681:1692,10] <- source_comp$Complexity_convolution

# add chain id (in this case we say that it's for chain 0)
all_data_w_source[1681:1692,3] <- rep(0,12)

# Now we take the next 12 rows and make them equal to the last 12 (that also contains source drawings). 
all_data_w_source[1693:1704,] <- all_data_w_source[1681:1692,]

# column 6 represents the chain id
all_data_w_source[1693:1704,3] <- rep(12,12) # chain 12

# We do this to all new batches of generation 0 source image rows. Not very tidy, but it works
all_data_w_source[1705:1716,] <- all_data_w_source[1681:1692,]
all_data_w_source[1705:1716,3] <- rep(13,12) # chain 13

all_data_w_source[1717:1728,] <- all_data_w_source[1681:1692,]
all_data_w_source[1717:1728,3] <- rep(14,12) # chain 14

all_data_w_source[1729:1740,] <- all_data_w_source[1681:1692,]
all_data_w_source[1729:1740,3] <- rep(15,12) # chain 15
                  
all_data_w_source[1741:1752,] <- all_data_w_source[1681:1692,]
all_data_w_source[1741:1752,3] <- rep(16,12)

all_data_w_source[1753:1764,] <- all_data_w_source[1681:1692,]
all_data_w_source[1753:1764,3] <- rep(17,12)

all_data_w_source[1765:1776,] <- all_data_w_source[1681:1692,]
all_data_w_source[1765:1776,3] <- rep(18,12)

all_data_w_source[1777:1788,] <- all_data_w_source[1681:1692,]
all_data_w_source[1777:1788,3] <- rep(19,12)

all_data_w_source[1789:1800,] <- all_data_w_source[1681:1692,]
all_data_w_source[1789:1800,3] <- rep(2,12)

all_data_w_source[1801:1812,] <- all_data_w_source[1681:1692,]
all_data_w_source[1801:1812,3] <- rep(20,12)

all_data_w_source[1813:1824,] <- all_data_w_source[1681:1692,]
all_data_w_source[1813:1824,3] <- rep(21,12)

all_data_w_source[1825:1836,] <- all_data_w_source[1681:1692,]
all_data_w_source[1825:1836,3] <- rep(22,12)

all_data_w_source[1837:1848,] <- all_data_w_source[1681:1692,]
all_data_w_source[1837:1848,3] <- rep(23,12)

all_data_w_source[1849:1860,] <- all_data_w_source[1681:1692,]
all_data_w_source[1849:1860,3] <- rep(3,12)

all_data_w_source[1861:1872,] <- all_data_w_source[1681:1692,]
all_data_w_source[1861:1872,3] <- rep(4,12)

all_data_w_source[1873:1884,] <- all_data_w_source[1681:1692,]
all_data_w_source[1873:1884,3] <- rep(5,12)

all_data_w_source[1885:1896,] <- all_data_w_source[1681:1692,]
all_data_w_source[1885:1896,3] <- rep(7,12)

all_data_w_source[1897:1908,] <- all_data_w_source[1681:1692,]
all_data_w_source[1897:1908,3] <- rep(8,12)

all_data_w_source[1909:1920,] <- all_data_w_source[1681:1692,]
all_data_w_source[1909:1920,3] <- rep(9,12)


# loop that inserts the correct subject ID, condition and source_image ID of the rest of the chain to each of the new generation 0 rows
for (i in 1:1920){ # i represents all the rows in the dataframe
  for (c in 0:23){ # c represents each of the 20 chains (their ID's goes from 0-23 because of missing chains)
    for (s in 1:12){ # s represents each of the 12 source images
      
      # below: if row i of column 3 (Chain) is equal to c, and row i of column 4 (generation) is 0 and column 6 (source image) is s
      if (all_data_w_source[i, 3] == c & all_data_w_source[i, 4] == 0 & all_data_w_source[i, 6] == s){
        
        # below: then take make the condition of row i (column 5) the same as the condition of row j and the same with subject ID (column 2)
        for (j in 1:1920){ # j represents the same row as i but exactly one generation above it
          if (all_data_w_source[j, 3] == c & all_data_w_source[j, 4] == 1 & all_data_w_source[j, 6] == s){
            all_data_w_source[i, 5] <- all_data_w_source[j, 5]
            all_data_w_source[i, 2] <- all_data_w_source[j, 2]
          }
        }
      }
    }
  }
}

```



Merge dataframes and save csv's
```{r}
# merge conventionality data with the dataset containing source images as gen 0 for each chain
all_data_w_all_conv_source <- merge(all_conventionality, all_data_w_source)

# write csv
write.csv(all_data_w_all_conv_source, "data/csv_files/all_data_w_all_conv_source.csv")
```




Plotting 

I chose to make barplots with the standard error as errorbars. Each plot shows the generation on the x-axis (from 0-7, 0 being the source images) and either the complexity or conventionality scores on the y-axis. For each generation, four bars represent each condition, color-coded as shown in the legend to the right of the plot.

```{r}

# Read data into R if you have cleaned the environment
#all_data_w_all_conv_source <- read.csv("data/csv_files/all_data_w_all_conv_source.csv")

# check classes of the variables of interest
class(all_data_w_all_conv_source$Condition)
class(all_data_w_all_conv_source$Generation)
class(all_data_w_all_conv_source$Conventionality)
class(all_data_w_all_conv_source$Complexity)

# change class for generation and condition to factor, and complexity and conventionality to numeric
all_data_w_all_conv_source$Condition <- as.factor(all_data_w_all_conv_source$Condition)
all_data_w_all_conv_source$Generation <- as.factor(all_data_w_all_conv_source$Generation)
all_data_w_all_conv_source$Conventionality <- as.numeric(all_data_w_all_conv_source$Conventionality)
all_data_w_all_conv_source$Complexity <- as.numeric(all_data_w_all_conv_source$Complexity)

# check range, mean and sd of conventionality and complexity
range(all_data_w_all_conv_source$Conventionality) # 0-10
mean(all_data_w_all_conv_source$Conventionality) # 4.33
sd(all_data_w_all_conv_source$Conventionality) # 2.80

range(all_data_w_all_conv_source$Complexity) # 4287-50067
mean(all_data_w_all_conv_source$Complexity) # 24641.25
sd(all_data_w_all_conv_source$Complexity) # 6399.96


# plotting complexity 
complexity_plot <- ggplot(all_data_w_all_conv_source, aes(Generation, Complexity, fill = Condition)) +
  geom_bar(fun.y = "mean", stat = "summary", position = "dodge") +
  ggtitle("Development of complexity") +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge")

complexity_plot

# plotting conventionality 
conventionality_plot <- ggplot(all_data_w_all_conv_source, aes(Generation, Conventionality, fill = Condition)) +
  geom_bar(fun.y = "mean", stat = "summary", position = "dodge") +
  ggtitle("Development of conventionality") +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge")

conventionality_plot

```


